A Constraint Solving Approach to Parikh Images of Regular Languages

### 2022-07-08 main.tex
That must have been a mistake
references

Parikh images play a central role in many automata-based algorithms,
for instance and notably in today's string solvers, which often have
to process constraints that combine regular language membership with
word length. To decide whether a simple formula like
$x \in \Language_1 \wedge y \in \Language_2 \wedge |x| > |y|$,
with string variables~$x, y$ and regular
languages~$\Language_1, \Language_2$, is satisfiable, it is
necessary to reason about the sets of word lengths induced by
$\Language_1, \Language_2$, which is a special case of the Parikh
image.  This required combined reasoning about strings and string
length has long been identified as a major bottleneck in string
solvers
\cite{DBLP:conf/cav/AbdullaACHRRS15,length-aware-solver,approximate-parikh,more}.
Other string solvers make use of Parikh automata~\cite{X}, and thus
Parikh images in the general case, to handle operations including
\verb!str.substr!  and \verb!str.at!, which comes at an even higher
price in terms of computational complexity~\cite{ostrich-plus}.

@@

It is possible to compute an existential Presburger formula describing
the Parikh image of any context-free language in linear
time~\cite{generate-parikh-image}; for the special case of regular
languages, this result was also stated in
\cite{muscholl-linear}. While theoretically elegant, this construction
has several disadvantages, often making it unpractical for integration
into algorithms. Firstly, the constructed Presburger formula contains
a linear number of existential quantifiers in the size of the
considered grammar, as well as complex Boolean structure, which is
needed to express the connectedness of sets of productions considered
in the construction. Eliminating those quantifiers to obtain a
quantifier-free representation of the Parikh image has exponential
complexity~\cite{X}, and is in practice often impossible in reasonable
time. Just solving the Parikh image membership problem is NP-complete,
as it corresponds to computing a satisfying assignment of the
existential Presburger formula, and taxing for solvers as
well~\cite{ostrich-plus}.

>>>>

Parikh images play a central role in many automata-based algorithms,
for instance and notably in today's string solvers, which often have
to process constraints that combine regular language membership with
word length. To decide whether a simple formula like
$x \in \Language_1 \wedge y \in \Language_2 \wedge |x| > |y|$,
with string variables~$x, y$ and regular
languages~$\Language_1, \Language_2$, is satisfiable, it is
necessary to reason about the sets of word lengths induced by
$\Language_1, \Language_2$, which is a special case of the Parikh
image.  This required combined reasoning about strings and string
length has long been identified as a major bottleneck in string
solvers
\cite{DBLP:conf/cav/AbdullaACHRRS15,length-aware-solver,approximate-parikh,DBLP:journals/corr/BerzishZG17}.
Other string solvers make use of Parikh automata~\cite{parikh-automata}, and thus
Parikh images in the general case, to handle operations including
\verb!str.substr!  and \verb!str.at!, which comes at an even higher
price in terms of computational complexity~\cite{ostrich-plus}.

@@

It is possible to compute an existential Presburger formula describing
the Parikh image of any context-free language in linear
time~\cite{generate-parikh-image}; for the special case of regular
languages, this result was also stated in
\cite{muscholl-linear}. While theoretically elegant, this construction
has several disadvantages, often making it unpractical for integration
into algorithms. Firstly, the constructed Presburger formula contains
a linear number of existential quantifiers in the size of the
considered grammar, as well as complex Boolean structure, which is
needed to express the connectedness of sets of productions considered
in the construction. Eliminating those quantifiers to obtain a
quantifier-free representation of the Parikh image has exponential
complexity~\cite{DBLP:conf/issac/Weispfenning97}, and is in practice often impossible in reasonable
time. Just solving the Parikh image membership problem is NP-complete,
as it corresponds to computing a satisfying assignment of the
existential Presburger formula, and taxing for solvers as
well~\cite{ostrich-plus}.
